---
title: "Introduction to Gratis"
author: "Bocong Zhao"
date: "17/07/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(fpp3)
library(parallel)
library(doParallel)
```

## About GRATIS

(1) The gratis package indicates generating time series with diverse and controllable characteristic, which can be used as an evaluation tool for tasks such as time series forecasting and classification. 

(2) Based on simulate time series data with mixture autoregressive model, gratis can coverage generalise time series and investigate the diversity in a time series feature space. 

(3) Furthermore, by tuning parameters of mixture autoregressive model, gratis can also efficiently generate new time series and controllable features. 

```{r setup}
# load package
library(gratis)
```

## Generate diverse time series

We use function **generate_ts()** to generate diverse time series

Our generation process use **distributions** instead of fixed parameter values in underlying models to allow generate diverse time series instances. The diversity of the generated time series should not rely on the parameter settings.

**Definitions**

Here are the definitions of parameter settings in function generate_ts():

|parameter settings | Definition|
|:----|:-----|
|n.ts | number of time series to be generated|
|freq |seasonal period of the time series to be generated|
|nComp|number of mixing components when simulating time series using MAR models|
|n    |length of the generated time series|

**Example**

Suppose we want to use MAR model to generate **3** time series from random parameter spaces. Each time series has **12** seasonal periods, **2** mixing components and the length **120**. 

```{r}
# Generate diverse time series
x <- generate_ts(n.ts = 3, freq = 12, nComp = 2, n = 120)
```

**Output**

(1) we can see **3** different time series be simulated, which are **N1**, **N2** and **N3**. In this example we use time series **N1** for further analysis.

(2) As required, there are **2** mixing components when simulating time series using MAR models, which are **pars1** and **pars2**

(3) Each component stands for different weight.

```{r}
# N1 time series 
data.matrix(x$N1$pars)
x$N1$pars 
```

**Plot time series**
```{r}
# plot N1 time series 
autoplot(x$N1$x)
```

## Generate mutiple seasonal time series

Time series can exhibit multiple seasonal pattern of different length, especially when series observed at a high frequency such as daily or hourly data.

We use function **generate_msts()** to generate mutiple seasonal time series.

**Definitions**

Here are the definitions of parameter settings in function generate_msts():

|parameter settings | Definition|
|:----|:-----|
|seasonal.periods | a vector of seasonal periods of the time series to be generated|
|nComp|number of mixing components when simulating time series using MAR models|
|n    |length of the generated time series|

**Example**

Suppose we want to use MAR model to generate a time series with **2** mixing components and the length **800** from random parameter spaces. Particularly, this time series has two seasonal periods **7** and **365**.

```{r}
# Generate mutiple seasonal time series
x <- generate_msts(seasonal.periods = c(7, 365), n = 800, nComp = 2)
```

**Plot time series**

```{r}
autoplot(x)
```

## Generate time series with controllable features

Time series analysis with particular focus may only interested in a certain area of feature space or a subset of features. 

Our function **generate_ts_with_target()** can efficiently generate time series with target features.  

The principle behind is that we use genetic algorithms to tune MAR parameters until the distance between *target feature vector* and *feature vector of a sample of time series simulated from MAR* is approximately equal to 0. 

**Definitions**

Here are the definitions of parameter settings in function generate_ts_with_target ():

|parameter settings | Definition|
|:----|:-----|
|n    |number of time series to be generated|
|ts.length|length of the time series to be generated|
|freq|frequency of the time series to be generated|
|seasonal|0 for non-seasonal data, 1 for single-seasonal data, and 2 for multiple seasonal data|
|features|a vector of function names|
|selected.features|selected features to be controlled|
|target|target feature values|
|parallel|An optional argument which allows to specify if the Genetic Algorithm should be run sequentially or in parallel|

**Example**

Suppose we want to use MAR model to generate **1 non-seasonal data** time series with frequency **1** and the length **60**. Particularly, this time series has two selected features, **entropy** and **trend** with target value between **0.6** to **0.9**

```{r}
x <- generate_ts_with_target(
  n = 1, ts.length = 60, freq = 1, seasonal = 0,
                        features = c('entropy', 'stl_features'),
                      selected.features = c('entropy', 'trend'),
                        target = c(0.6, 0.9))
```

**Plot time series**

```{r}
autoplot(x)
```

